### 2021-03-23 thesis/intro/main.tex

\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}

\usepackage{float}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{tabu}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{subcaption}
%\captionsetup{compatibility=false}

%\usepackage{apacite}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Full name of the applicant: Aspirant [ASP]
SEMAPHORE Application ID: 34791278}
\author{Energy-Optimal Configurations for HPC Applications - Vitor Ramos Gomes da Silva}
\date{\today}

\begin{document}
\maketitle

\input{paper}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}