### 2021-03-23 thesis/pascalanalyzer/DataStructure2.tex

\documentclass[11pt]{article}

    \usepackage[breakable]{tcolorbox}
    \usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)

    \usepackage{iftex}
    \ifPDFTeX
    	\usepackage[T1]{fontenc}
    	\usepackage{mathpazo}
    \else
    	\usepackage{fontspec}
    \fi

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % Maintain compatibility with old templates. Remove in nbconvert 6.0
    \let\Oldincludegraphics\includegraphics
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionFormat{nocaption}{}
    \captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}

    \usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
    \adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}
    \usepackage{float}
    \floatplacement{figure}{H} % forces figures to be placed at the correct location
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range
    \makeatletter % fix for grffile with XeLaTeX
    \def\Gread@@xetex#1{%
      \IfFileExists{"\Gin@base".bb}%
      {\Gread@eps{\Gin@base.bb}}%
      {\Gread@@xetex@aux#1}%
    }
    \makeatother

    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    % The default LaTeX title has an obnoxious amount of whitespace. By default,
    % titling removes some of it. It also provides customization options.
    \usepackage{titling}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines
    \usepackage{mathrsfs}

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}
    \definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
    \definecolor{ansi-default-inverse-bg}{HTML}{000000}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatibility definitions
    \def\gt{>}
    \def\lt{<}
    \let\Oldtex\TeX
    \let\Oldlatex\LaTeX
    \renewcommand{\TeX}{\textrm{\Oldtex}}
    \renewcommand{\LaTeX}{\textrm{\Oldlatex}}
    % Document parameters
    % Document title
    \title{DataStructure2}

% Pygments definitions
\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % For linebreaks inside Verbatim environment from package fancyvrb. 
    \makeatletter
        \newbox\Wrappedcontinuationbox 
        \newbox\Wrappedvisiblespacebox 
        \newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}} 
        \newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}} 
        \newcommand*\Wrappedcontinuationindent {3ex } 
        \newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox} 
        % Take advantage of the already applied Pygments mark-up to insert 
        % potential linebreaks for TeX processing. 
        %        {, <, #, %, $, ' and ": go to next line. 
        %        _, }, ^, &, >, - and ~: stay at end of broken line. 
        % Use of \textquotesingle for straight quote. 
        \newcommand*\Wrappedbreaksatspecials {% 
            \def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}% 
            \def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}% 
            \def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}% 
            \def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}% 
            \def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}% 
            \def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}% 
            \def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}% 
            \def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}% 
            \def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}% 
            \def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}% 
            \def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}% 
            \def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}% 
            \def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}% 
            \def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}% 
        } 
        % Some characters . , ; ? ! / are not pygmentized. 
        % This macro makes them "active" and they will insert potential linebreaks 
        \newcommand*\Wrappedbreaksatpunct {% 
            \lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}% 
            \lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}% 
            \lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}% 
            \lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}% 
            \lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}% 
            \lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}% 
            \lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}% 
            \catcode`\.\active
            \catcode`\,\active 
            \catcode`\;\active
            \catcode`\:\active
            \catcode`\?\active
            \catcode`\!\active
            \catcode`\/\active 
            \lccode`\~`\~ 	
        }
    \makeatother

    \let\OriginalVerbatim=\Verbatim
    \makeatletter
    \renewcommand{\Verbatim}[1][1]{%
        %\parskip\z@skip
        \sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
        \sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
        \def\FancyVerbFormatLine ##1{\hsize\linewidth
            \vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
                \doublehyphendemerits\z@\finalhyphendemerits\z@
                \strut ##1\strut}%
        }%
        % If the linebreak is at a space, the latter will be displayed as visible
        % space at end of first line, and a continuation symbol starts next line.
        % Stretch/shrink are however usually zero for typewriter font.
        \def\FV@Space {%
            \nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
            \discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
            {\kern\fontdimen2\font}%
        }%

        % Allow breaks at special characters using \PYG... macros.
        \Wrappedbreaksatspecials
        % Breaks at punctuation characters . , ; ? ! and / need catcode=\active 	
        \OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
    }
    \makeatother

    % Exact colors from NB
    \definecolor{incolor}{HTML}{303F9F}
    \definecolor{outcolor}{HTML}{D84315}
    \definecolor{cellborder}{HTML}{CFCFCF}
    \definecolor{cellbackground}{HTML}{F7F7F7}

    % prompt
    \makeatletter
    \newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
    \makeatother
    \newcommand{\prompt}[4]{
        \ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}
    }

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

    \maketitle

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{collections} \PY{k+kn}{import} \PY{n}{defaultdict}
\PY{k+kn}{import} \PY{n+nn}{json}
\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}

\PY{k}{def} \PY{n+nf}{multidict}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{:}
    \PY{k}{if} \PY{n}{n} \PY{o}{==} \PY{l+m+mi}{0}\PY{p}{:} \PY{k}{return} \PY{n}{dtype}
    \PY{k}{return} \PY{n}{defaultdict}\PY{p}{(}\PY{k}{lambda}\PY{p}{:}\PY{n}{multidict}\PY{p}{(}\PY{n}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,}\PY{n}{dtype}\PY{p}{)}\PY{p}{)}

\PY{n}{nested\PYZus{}dict} \PY{o}{=} \PY{k}{lambda}\PY{p}{:} \PY{n}{defaultdict}\PY{p}{(}\PY{n}{nested\PYZus{}dict}\PY{p}{)}

\PY{n}{freqs}\PY{o}{=} \PY{p}{[}\PY{l+m+mf}{1.2}\PY{p}{,}\PY{l+m+mf}{2.2}\PY{p}{]}
\PY{n}{cores}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{32}\PY{p}{]}
\PY{n}{inps}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{]}
\PY{n}{regions}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{measures} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
\PY{c+c1}{\PYZsh{} measures= defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:\PYZob{}\PYZcb{})))}

\PY{c+c1}{\PYZsh{} collect every n milliseconds + event}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{o}{=} \PY{p}{\PYZob{}}
                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{200}\PY{p}{\PYZcb{}}\PY{p}{,} \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.2}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{220}\PY{p}{\PYZcb{}}\PY{p}{]}\PY{p}{,}
                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{total\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{10}\PY{p}{,}
                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{total\PYZus{}energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{10}
                    \PY{p}{\PYZcb{}}
                \PY{c+c1}{\PYZsh{} measures[e][f][c][i]= \PYZob{}}
                \PY{c+c1}{\PYZsh{}     \PYZdq{}sensor\PYZdq{}:[\PYZob{}\PYZdq{}timestamp\PYZdq{}:0.1,\PYZdq{}data\PYZdq{}:200\PYZcb{}, \PYZob{}\PYZdq{}timestamp\PYZdq{}:0.2,\PYZdq{}data\PYZdq{}:220\PYZcb{}],}
                \PY{c+c1}{\PYZsh{}     \PYZdq{}total\PYZus{}time\PYZdq{}:10,}
                \PY{c+c1}{\PYZsh{}     \PYZdq{}total\PYZus{}energy\PYZdq{}:10}
                \PY{c+c1}{\PYZsh{}     \PYZcb{}}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{o}{=} \PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.32}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{150}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{]}
                    \PY{c+c1}{\PYZsh{} measures[e][r][f][c][i]= [\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:\PYZob{}\PYZdq{}timestamp\PYZdq{}:0.32,\PYZdq{}data\PYZdq{}:150\PYZcb{}\PYZcb{},]}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,} \PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{
    "1.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \}
    \},
    "2.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \}
    \}
\}

    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{measures} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
\PY{c+c1}{\PYZsh{} measures= defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:\PYZob{}\PYZcb{}))))}
\PY{c+c1}{\PYZsh{} collect every event}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{o}{=} \PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{5}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{]}
                    \PY{c+c1}{\PYZsh{} measures[e][r][f][c][i]= [\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:5\PYZcb{},]}
                    \PY{c+c1}{\PYZsh{} measures[\PYZdq{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZdq{}\PYZpc{}(f,c,i,e,r)]=[\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:5\PYZcb{},]}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{
    "1.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \}
    \},
    "2.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \}
    \}
\}

    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} difficulty to transform this into a xarray, dataframe, numpy}
\PY{c+c1}{\PYZsh{} performance of execution operations on each format}

\PY{n}{measures}\PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{e}\PY{p}{,}\PY{n}{r}\PY{p}{)}\PY{p}{]}\PY{o}{=}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{5}\PY{p}{\PYZcb{}}\PY{p}{]}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}
    \PY{n+nb}{map}\PY{p}{(}\PY{k}{lambda} \PY{n}{x}\PY{p}{:} \PY{n}{x}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{+}\PY{p}{[}\PY{n}{measures}\PY{p}{[}\PY{n}{x}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{]}\PY{p}{,}\PY{n}{measures}\PY{p}{)}\PY{p}{,}
    \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{
    "1.2-16-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-16-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-16-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-16-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ]
\}

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/vitor/anaconda3/lib/python3.7/site-packages/dask/config.py:168:
YAMLLoadWarning: calling yaml.load() without Loader={\ldots} is deprecated, as the
default Loader is unsafe. Please read https://msg.pyyaml.org/load for full
details.
  data = yaml.load(f.read()) or \{\}
/home/vitor/anaconda3/lib/python3.7/site-packages/distributed/config.py:20:
YAMLLoadWarning: calling yaml.load() without Loader={\ldots} is deprecated, as the
default Loader is unsafe. Please read https://msg.pyyaml.org/load for full
details.
  defaults = yaml.load(f)
    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
                             sensor
freq cores inputs execs reg
1.2  16    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
     32    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
2.2  16    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
     32    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 1, freq: 2, inputs: 2, reg: 2)
Coordinates:
  * freq     (freq) object '1.2' '2.2'
  * cores    (cores) object '16' '32'
  * inputs   (inputs) object '2' '5'
  * execs    (execs) object '0'
  * reg      (reg) object '2' '4'
Data variables:
    sensor   (freq, cores, inputs, execs, reg) int64 5 5 5 5 5 5 {\ldots} 5 5 5 5 5 5
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} difficulty to transform this into a xarray, dataframe, numpy}
\PY{c+c1}{\PYZsh{} performance of execution operations on each format}

\PY{n}{measures}\PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{e}\PY{p}{,}\PY{n}{r}\PY{p}{)}\PY{p}{]}\PY{o}{=}\PY{p}{[}
                        \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{,}
                        \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{,}
                        \PY{p}{[}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{]}
                        \PY{p}{]}

\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}
    \PY{n+nb}{map}\PY{p}{(}\PY{k}{lambda} \PY{n}{x}\PY{p}{:} \PY{n}{x}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{+}\PY{n}{measures}\PY{p}{[}\PY{n}{x}\PY{p}{]}\PY{p}{,}\PY{n}{measures}\PY{p}{)}\PY{p}{,}
    \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
                                 start       stop     sensor
freq cores inputs execs reg
1.2  16    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
     32    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
2.2  16    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
     32    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 1, freq: 2, inputs: 2, reg: 2)
Coordinates:
  * freq     (freq) object '1.2' '2.2'
  * cores    (cores) object '16' '32'
  * inputs   (inputs) object '2' '5'
  * execs    (execs) object '0'
  * reg      (reg) object '2' '4'
Data variables:
    start    (freq, cores, inputs, execs, reg) object [0, 1, 3] {\ldots} [0, 1, 3]
    stop     (freq, cores, inputs, execs, reg) object [2, 3, 4] {\ldots} [2, 3, 4]
    sensor   (freq, cores, inputs, execs, reg) object [5, 4, 5] {\ldots} [5, 4, 5]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}

                \PY{n}{aux}\PY{o}{=} \PY{p}{[}\PY{p}{]}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{c+c1}{\PYZsh{} start\PYZus{}time stop\PYZus{}time sensor}
                    \PY{n}{aux}\PY{o}{.}\PY{n}{append}\PY{p}{(}
                        \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{region}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{n}{r}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{)}
                    \PY{n}{aux}\PY{o}{.}\PY{n}{append}\PY{p}{(}
                        \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{region}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{n}{r}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{)}

                \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{aux}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{)}
\PY{n}{idxs}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{MultiIndex}\PY{o}{.}\PY{n}{from\PYZus{}product}\PY{p}{(}\PY{p}{[}\PY{n}{execs}\PY{p}{,}\PY{n}{freqs}\PY{p}{,}\PY{n}{cores}\PY{p}{,}\PY{n}{inps}\PY{p}{]}\PY{p}{,}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{df}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
                                                                         0  \textbackslash{}
freq cores inputs execs
1    1.2   16     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
     2.2   16     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}

                                                                         1  \textbackslash{}
freq cores inputs execs
1    1.2   16     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
     2.2   16     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}

                                                                         2  \textbackslash{}
freq cores inputs execs
1    1.2   16     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
     2.2   16     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}

                                                                         3
freq cores inputs execs
1    1.2   16     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
     2.2   16     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 2, freq: 1, inputs: 2)
Coordinates:
  * freq     (freq) int64 1
  * cores    (cores) float64 1.2 2.2
  * inputs   (inputs) int64 16 32
  * execs    (execs) int64 2 5
Data variables:
    0        (freq, cores, inputs, execs) object \{'region': 2, 'start\_time': 0,
'stop\_time': 2, 'sensor': \{\}\} {\ldots} \{'region': 2, 'start\_time': 0, 'stop\_time': 2,
'sensor': \{\}\}
    1        (freq, cores, inputs, execs) object \{'region': 2, 'start\_time': 3,
'stop\_time': 4, 'sensor': \{\}\} {\ldots} \{'region': 2, 'start\_time': 3, 'stop\_time': 4,
'sensor': \{\}\}
    2        (freq, cores, inputs, execs) object \{'region': 4, 'start\_time': 0,
'stop\_time': 2, 'sensor': \{\}\} {\ldots} \{'region': 4, 'start\_time': 0, 'stop\_time': 2,
'sensor': \{\}\}
    3        (freq, cores, inputs, execs) object \{'region': 4, 'start\_time': 3,
'stop\_time': 4, 'sensor': \{\}\} {\ldots} \{'region': 4, 'start\_time': 3, 'stop\_time': 4,
'sensor': \{\}\}
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[}\PY{n}{e}\PY{p}{,}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{r}\PY{p}{,}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados1}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{]}\PY{p}{]}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
                                                     dados
exec freq cores inputs reg
1    1.2  16    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
          32    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
     2.2  16    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
          32    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
<xarray.Dataset>
Dimensions:  (cores: 2, exec: 1, freq: 2, inputs: 2, reg: 2)
Coordinates:
  * exec     (exec) int64 1
  * freq     (freq) float64 1.2 2.2
  * cores    (cores) int64 16 32
  * inputs   (inputs) int64 2 5
  * reg      (reg) int64 2 4
Data variables:
    dados    (exec, freq, cores, inputs, reg) object [\{'dados1': 0\}, \{'dados2':
0\}] {\ldots} [\{'dados1': 0\}, \{'dados2': 0\}]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{]}
\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}

                \PY{n}{aux}\PY{o}{=} \PY{n}{defaultdict}\PY{p}{(}\PY{k}{lambda}\PY{p}{:}\PY{p}{[}\PY{p}{[}\PY{p}{]}\PY{p}{,}\PY{p}{[}\PY{p}{]}\PY{p}{,}\PY{p}{[}\PY{p}{]}\PY{p}{]}\PY{p}{)}
                \PY{k}{for} \PY{n}{rx} \PY{o+ow}{in} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{:}
                    \PY{c+c1}{\PYZsh{} start\PYZus{}time stop\PYZus{}time sensor}
                    \PY{c+c1}{\PYZsh{} aux[rx].append(}
                    \PY{c+c1}{\PYZsh{}     \PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:2,\PYZdq{}sensor\PYZdq{}:\PYZob{}\PYZcb{}\PYZcb{})}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{)}

                \PY{n}{reg\PYZus{}k}\PY{o}{=} \PY{n}{aux}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{aux}\PY{p}{:}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
                    \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{c+c1}{\PYZsh{} df[\PYZdq{}start\PYZdq{}]= df[\PYZdq{}start\PYZdq{}].apply(lambda x: np.array(x))}
\PY{c+c1}{\PYZsh{} df[\PYZdq{}stop\PYZdq{}]= df[\PYZdq{}stop\PYZdq{}].apply(lambda x: np.array(x))}
\PY{n}{idxs}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{MultiIndex}\PY{o}{.}\PY{n}{from\PYZus{}product}\PY{p}{(}\PY{p}{[}\PY{n}{execs}\PY{p}{,}\PY{n}{freqs}\PY{p}{,}\PY{n}{cores}\PY{p}{,}\PY{n}{inps}\PY{p}{,}\PY{n}{reg\PYZus{}k}\PY{p}{]}\PY{p}{,}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{regs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{df}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 2, freq: 2, inputs: 2, regs: 2)
Coordinates:
  * execs    (execs) int64 1 2
  * freq     (freq) float64 1.2 2.2
  * cores    (cores) int64 16 32
  * inputs   (inputs) int64 2 5
  * regs     (regs) int64 2 4
Data variables:
    start    (execs, freq, cores, inputs, regs) object [0 0] [0 0] {\ldots} [0 0]
    stop     (execs, freq, cores, inputs, regs) object [1 1] [1 1] {\ldots} [1 1]
    sensor   (execs, freq, cores, inputs, regs) object [\{\} \{\}] {\ldots} [\{\} \{\}]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{data}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
                \PY{k}{for} \PY{n}{x} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{randint}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{)}\PY{p}{)}\PY{p}{:}
                    \PY{n}{data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[}\PY{n}{e}\PY{p}{,}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{x}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
\PY{n}{dfx}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{data}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{core}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ent}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{dfx}\PY{o}{=} \PY{n}{dfx}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{core}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ent}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{dfx}
\PY{c+c1}{\PYZsh{} xrx= dfx.to\PYZus{}xarray()}
\PY{c+c1}{\PYZsh{} xrx}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
                        data
exec freq core inp ent
1    1.2  16   2   0       1
                   1       1
                   2       1
                   3       1
               5   0       1
                   1       1
                   2       1
          32   2   0       1
               5   0       1
     2.2  16   2   0       1
                   1       1
                   2       1
               5   0       1
                   1       1
          32   2   0       1
               5   0       1
                   1       1
2    1.2  16   2   0       1
                   1       1
               5   0       1
                   1       1
          32   2   0       1
                   1       1
                   2       1
               5   0       1
                   1       1
                   2       1
     2.2  16   2   0       1
                   1       1
                   2       1
                   3       1
               5   0       1
          32   2   0       1
               5   0       1
\end{Verbatim}
\end{tcolorbox}

    \hypertarget{benchmarks}{%
\section{Benchmarks}\label{benchmarks}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{itertools} \PY{k+kn}{import} \PY{n}{product}
\PY{k+kn}{import} \PY{n+nn}{random}
\PY{k+kn}{import} \PY{n+nn}{string}
\PY{k+kn}{import} \PY{n+nn}{sys}
\PY{k+kn}{from} \PY{n+nn}{numpy} \PY{k+kn}{import} \PY{n}{prod}
\PY{k+kn}{import} \PY{n+nn}{pickle}
\PY{k+kn}{import} \PY{n+nn}{gzip}
\PY{k+kn}{import} \PY{n+nn}{os}
\PY{k+kn}{import} \PY{n+nn}{gc}

\PY{k}{def} \PY{n+nf}{si\PYZus{}format}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n}{precision}\PY{o}{=} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{:}
    \PY{n}{sufx}\PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{M}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{1e6}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{K}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{1e3}\PY{p}{\PYZcb{}}
    \PY{n}{sufix}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
    \PY{k}{for} \PY{n}{k}\PY{p}{,}\PY{n}{v} \PY{o+ow}{in} \PY{n}{sufx}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}\PY{p}{:}
        \PY{k}{if} \PY{n}{n} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{n}{v}\PY{p}{:}
            \PY{n}{n}\PY{o}{/}\PY{o}{=}\PY{n}{v}
            \PY{n}{sufix}\PY{o}{=} \PY{n}{k}
            \PY{k}{break}
    \PY{k}{return} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{:.}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{f\PYZcb{} }\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{b}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{n}\PY{p}{,}\PY{n}{precision}\PY{p}{,}\PY{n}{sufix}\PY{p}{)}

\PY{k}{def} \PY{n+nf}{get\PYZus{}size}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{n}{seen}\PY{o}{=}\PY{k+kc}{None}\PY{p}{)}\PY{p}{:}
    \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Recursively finds size of objects\PYZdq{}\PYZdq{}\PYZdq{}}
    \PY{n}{size} \PY{o}{=} \PY{n}{sys}\PY{o}{.}\PY{n}{getsizeof}\PY{p}{(}\PY{n}{obj}\PY{p}{)}
    \PY{k}{if} \PY{n}{seen} \PY{o+ow}{is} \PY{k+kc}{None}\PY{p}{:}
        \PY{n}{seen} \PY{o}{=} \PY{n+nb}{set}\PY{p}{(}\PY{p}{)}
    \PY{n}{obj\PYZus{}id} \PY{o}{=} \PY{n+nb}{id}\PY{p}{(}\PY{n}{obj}\PY{p}{)}
    \PY{k}{if} \PY{n}{obj\PYZus{}id} \PY{o+ow}{in} \PY{n}{seen}\PY{p}{:}
        \PY{k}{return} \PY{l+m+mi}{0}
    \PY{c+c1}{\PYZsh{} Important mark as seen *before* entering recursion to gracefully handle}
    \PY{c+c1}{\PYZsh{} self\PYZhy{}referential objects}
    \PY{n}{seen}\PY{o}{.}\PY{n}{add}\PY{p}{(}\PY{n}{obj\PYZus{}id}\PY{p}{)}
    \PY{k}{if} \PY{n+nb}{isinstance}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{n+nb}{dict}\PY{p}{)}\PY{p}{:}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{v}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{obj}\PY{o}{.}\PY{n}{values}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{)}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{k}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{obj}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{)}
    \PY{k}{elif} \PY{n+nb}{hasattr}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZus{}\PYZus{}dict\PYZus{}\PYZus{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}\PY{p}{:}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{obj}\PY{o}{.}\PY{n+nv+vm}{\PYZus{}\PYZus{}dict\PYZus{}\PYZus{}}\PY{p}{,} \PY{n}{seen}\PY{p}{)}
    \PY{k}{elif} \PY{n+nb}{hasattr}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZus{}\PYZus{}iter\PYZus{}\PYZus{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o+ow}{and} \PY{o+ow}{not} \PY{n+nb}{isinstance}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{p}{(}\PY{n+nb}{str}\PY{p}{,} \PY{n+nb}{bytes}\PY{p}{,} \PY{n+nb}{bytearray}\PY{p}{)}\PY{p}{)}\PY{p}{:}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{obj}\PY{p}{]}\PY{p}{)}
    \PY{k}{return} \PY{n}{size}

\PY{c+c1}{\PYZsh{} cores freq input exec reg=}
\PY{k}{def} \PY{n+nf}{bench}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{,} \PY{n}{nbytes}\PY{p}{)}\PY{p}{:}
    \PY{n}{vx}\PY{o}{=} \PY{p}{[}\PY{p}{]}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Keys:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{prod}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ Bytes:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{c+c1}{\PYZsh{}     vx= [[ \PYZdq{}var\PYZob{}\PYZcb{}\PYZus{}\PYZob{}\PYZcb{}\PYZdq{}.format(i, j) for j in range(vs) ] for i, vs in enumerate(v\PYZus{}sizes)]}
    \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{vs} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{:}
        \PY{n}{vx}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{var}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{vs}\PY{p}{)} \PY{p}{]}\PY{p}{)}

    \PY{n}{multi\PYZus{}dict} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
    \PY{n}{normal\PYZus{}dict} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{vx}\PY{p}{)}\PY{p}{:}
        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}

        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
        \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{o}{=} \PY{n}{data}

        \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}\PY{o}{=} \PY{n}{data}

    \PY{n}{s1}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{RAM       }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{w+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{w+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}\PY{p}{)}

    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK json }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{pickle}\PY{o}{.}\PY{n}{dump}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{loads}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{n}{tf}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{pickle}\PY{o}{.}\PY{n}{dump}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{,} \PY{n}{tf}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}

    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK pkl  }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}

    \PY{k}{with} \PY{n}{gzip}\PY{o}{.}\PY{n}{GzipFile}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{o}{.}\PY{n}{encode}\PY{p}{(}\PY{p}{)}\PY{p}{)}

    \PY{k}{with} \PY{n}{gzip}\PY{o}{.}\PY{n}{GzipFile}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}\PY{o}{.}\PY{n}{encode}\PY{p}{(}\PY{p}{)}\PY{p}{)}

    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK gz  }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2}\PY{p}{)}

    \PY{n}{gc}\PY{o}{.}\PY{n}{collect}\PY{p}{(}\PY{p}{)}

\PY{c+c1}{\PYZsh{}bench([2,3,2],10)}
\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],10)}
\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],1e3)}
\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],100e3)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e1}\PY{p}{)}
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e2}\PY{p}{)}
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e3}\PY{p}{)}
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 24.00 Kb  Bytes: 10.00 b

RAM        2.57 Mb 4.73 Mb 0.542656059372869
DISK json  616.52 Kb 1.26 Mb 0.49046778042959427
DISK pkl   380.99 Kb 1.21 Mb 0.31507013577280013
DISK gz   221.87 Kb 270.49 Kb 0.8202447410255462

Keys: 24.00 Kb  Bytes: 100.00 b

RAM        4.73 Mb 6.89 Mb 0.6860778510924217
DISK json  2.78 Mb 3.42 Mb 0.8125601404741001
DISK pkl   2.54 Mb 3.37 Mb 0.7541981461470956
DISK gz   1.82 Mb 1.87 Mb 0.9686695423189734

Keys: 24.00 Kb  Bytes: 1.00 Kb

RAM        26.33 Mb 28.49 Mb 0.9240995518926407
DISK json  24.38 Mb 25.02 Mb 0.974398129272095
DISK pkl   24.22 Mb 25.04 Mb 0.9669301631449064
DISK gz   17.58 Mb 17.64 Mb 0.9962784816626885

Keys: 24.00 Kb  Bytes: 10.00 Kb

RAM        242.33 Mb 244.49 Mb 0.9911560500830847
DISK json  240.38 Mb 241.02 Mb 0.9973425857927034
DISK pkl   240.24 Mb 241.07 Mb 0.996564850797469
DISK gz   175.11 Mb 175.07 Mb 1.0001946831608381
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{32}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 320.00 Kb  Bytes: 1.00 b

RAM        13.38 Mb 37.27 Mb 0.3588984160511397
DISK json  5.23 Mb 13.98 Mb 0.3738659513590844
DISK pkl   1.49 Mb 12.72 Mb 0.1175192426066018
DISK gz   537.40 Kb 1.24 Mb 0.4322204295799909
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{import} \PY{n+nn}{time}

\PY{k}{def} \PY{n+nf}{bench2}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{,} \PY{n}{nbytes}\PY{p}{)}\PY{p}{:}
    \PY{n}{vx}\PY{o}{=} \PY{p}{[}\PY{p}{]}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Keys:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{prod}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ Bytes:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{vs} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{:}
        \PY{n}{vx}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{var}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{vs}\PY{p}{)} \PY{p}{]}\PY{p}{)}

    \PY{n}{multi\PYZus{}dict} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
    \PY{n}{normal\PYZus{}dict} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
    \PY{n}{all\PYZus{}comb}\PY{o}{=} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{vx}\PY{p}{)}
    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}

        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
        \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{o}{=} \PY{n}{data}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s1}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}

    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}
        \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}\PY{o}{=} \PY{n}{data}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Creation time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{s1}\PY{p}{,} \PY{n}{s2}\PY{p}{)}

    \PY{c+c1}{\PYZsh{}multi\PYZus{}dict= json.loads(json.dumps(multi\PYZus{}dict))}
    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
        \PY{n}{z}\PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s1}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}

    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{z}\PY{o}{=} \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Acess time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{s1}\PY{p}{,} \PY{n}{s2}\PY{p}{)}

    \PY{n}{gc}\PY{o}{.}\PY{n}{collect}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e1}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e2}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e3}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 24.00 Kb  Bytes: 10.00 b

Creation time 0.12965726852416992 2.384185791015625e-07
Acess time 4.76837158203125e-07 2.384185791015625e-07

Keys: 24.00 Kb  Bytes: 100.00 b

Creation time 0.6197633743286133 4.76837158203125e-07
Acess time 2.384185791015625e-07 4.76837158203125e-07

Keys: 24.00 Kb  Bytes: 1.00 Kb

Creation time 5.764535903930664 2.384185791015625e-07
Acess time 7.152557373046875e-07 2.384185791015625e-07

Keys: 24.00 Kb  Bytes: 10.00 Kb

Creation time 56.628950119018555 7.152557373046875e-07
Acess time 4.76837158203125e-07 0.0
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{*}\PY{l+m+mi}{1000}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{*}\PY{l+m+mi}{1000}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 1.00 b  Bytes: 10.00 Kb

RAM        314.94 Kb 19.23 Kb 16.38003848754356
DISK json  23.89 Kb 18.90 Kb 1.2643276710588982
DISK pkl   23.91 Kb 18.92 Kb 1.263903573694227
DISK gz   10.34 Kb 10.14 Kb 1.020122312093115

Keys: 1.00 b  Bytes: 10.00 Kb

Creation time 0.002998828887939453 4.76837158203125e-07
Acess time 7.152557373046875e-07 2.384185791015625e-07
    \end{Verbatim}

    % Add a bibliography block to the postdoc

\end{document}