### 2021-03-23 thesis/pascalanalyzer/arango.tex

\documentclass[11pt]{article}

    \usepackage[breakable]{tcolorbox}
    \usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)

    \usepackage{iftex}
    \ifPDFTeX
    	\usepackage[T1]{fontenc}
    	\usepackage{mathpazo}
    \else
    	\usepackage{fontspec}
    \fi

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % Maintain compatibility with old templates. Remove in nbconvert 6.0
    \let\Oldincludegraphics\includegraphics
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionFormat{nocaption}{}
    \captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}

    \usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
    \adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}
    \usepackage{float}
    \floatplacement{figure}{H} % forces figures to be placed at the correct location
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range
    \makeatletter % fix for grffile with XeLaTeX
    \def\Gread@@xetex#1{%
      \IfFileExists{"\Gin@base".bb}%
      {\Gread@eps{\Gin@base.bb}}%
      {\Gread@@xetex@aux#1}%
    }
    \makeatother

    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    % The default LaTeX title has an obnoxious amount of whitespace. By default,
    % titling removes some of it. It also provides customization options.
    \usepackage{titling}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines
    \usepackage{mathrsfs}

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}
    \definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
    \definecolor{ansi-default-inverse-bg}{HTML}{000000}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatibility definitions
    \def\gt{>}
    \def\lt{<}
    \let\Oldtex\TeX
    \let\Oldlatex\LaTeX
    \renewcommand{\TeX}{\textrm{\Oldtex}}
    \renewcommand{\LaTeX}{\textrm{\Oldlatex}}
    % Document parameters
    % Document title
    \title{arango}

% Pygments definitions
\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % For linebreaks inside Verbatim environment from package fancyvrb. 
    \makeatletter
        \newbox\Wrappedcontinuationbox 
        \newbox\Wrappedvisiblespacebox 
        \newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}} 
        \newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}} 
        \newcommand*\Wrappedcontinuationindent {3ex } 
        \newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox} 
        % Take advantage of the already applied Pygments mark-up to insert 
        % potential linebreaks for TeX processing. 
        %        {, <, #, %, $, ' and ": go to next line. 
        %        _, }, ^, &, >, - and ~: stay at end of broken line. 
        % Use of \textquotesingle for straight quote. 
        \newcommand*\Wrappedbreaksatspecials {% 
            \def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}% 
            \def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}% 
            \def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}% 
            \def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}% 
            \def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}% 
            \def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}% 
            \def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}% 
            \def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}% 
            \def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}% 
            \def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}% 
            \def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}% 
            \def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}% 
            \def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}% 
            \def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}% 
        } 
        % Some characters . , ; ? ! / are not pygmentized. 
        % This macro makes them "active" and they will insert potential linebreaks 
        \newcommand*\Wrappedbreaksatpunct {% 
            \lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}% 
            \lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}% 
            \lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}% 
            \lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}% 
            \lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}% 
            \lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}% 
            \lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}% 
            \catcode`\.\active
            \catcode`\,\active 
            \catcode`\;\active
            \catcode`\:\active
            \catcode`\?\active
            \catcode`\!\active
            \catcode`\/\active 
            \lccode`\~`\~ 	
        }
    \makeatother

    \let\OriginalVerbatim=\Verbatim
    \makeatletter
    \renewcommand{\Verbatim}[1][1]{%
        %\parskip\z@skip
        \sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
        \sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
        \def\FancyVerbFormatLine ##1{\hsize\linewidth
            \vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
                \doublehyphendemerits\z@\finalhyphendemerits\z@
                \strut ##1\strut}%
        }%
        % If the linebreak is at a space, the latter will be displayed as visible
        % space at end of first line, and a continuation symbol starts next line.
        % Stretch/shrink are however usually zero for typewriter font.
        \def\FV@Space {%
            \nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
            \discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
            {\kern\fontdimen2\font}%
        }%

        % Allow breaks at special characters using \PYG... macros.
        \Wrappedbreaksatspecials
        % Breaks at punctuation characters . , ; ? ! and / need catcode=\active 	
        \OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
    }
    \makeatother

    % Exact colors from NB
    \definecolor{incolor}{HTML}{303F9F}
    \definecolor{outcolor}{HTML}{D84315}
    \definecolor{cellborder}{HTML}{CFCFCF}
    \definecolor{cellbackground}{HTML}{F7F7F7}

    % prompt
    \makeatletter
    \newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
    \makeatother
    \newcommand{\prompt}[4]{
        \ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}
    }

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

    \maketitle

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{138}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{arango} \PY{k+kn}{import} \PY{n}{ArangoClient}
\PY{k+kn}{import} \PY{n+nn}{os}\PY{o}{,} \PY{n+nn}{json}
\PY{c+c1}{\PYZsh{} Initialize the ArangoDB client.}
\PY{n}{client} \PY{o}{=} \PY{n}{ArangoClient}\PY{p}{(}\PY{n}{hosts}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{http://172.17.0.2:8529}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{sys\PYZus{}db} \PY{o}{=} \PY{n}{client}\PY{o}{.}\PY{n}{db}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}system}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{username}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{root}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{password}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{oN1WiCr2sm4m1H4g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{k}{if} \PY{o+ow}{not} \PY{n}{sys\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}database}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{sys\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}database}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{energy\PYZus{}db}\PY{o}{=} \PY{n}{client}\PY{o}{.}\PY{n}{db}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{username}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{root}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{password}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{oN1WiCr2sm4m1H4g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{k}{if} \PY{o+ow}{not} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{run\PYZus{}col}\PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{k}{if} \PY{o+ow}{not} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{model\PYZus{}col}\PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{k}{if} \PY{o+ow}{not} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}graph}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}graph}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{md\PYZus{}graph}\PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{graph}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{k}{if} \PY{o+ow}{not} \PY{n}{md\PYZus{}graph}\PY{o}{.}\PY{n}{has\PYZus{}edge\PYZus{}definition}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{md\PYZus{}graph}\PY{o}{.}\PY{n}{create\PYZus{}edge\PYZus{}definition}\PY{p}{(}
        \PY{n}{edge\PYZus{}collection}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
        \PY{n}{from\PYZus{}vertex\PYZus{}collections}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
        \PY{n}{to\PYZus{}vertex\PYZus{}collections}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
    \PY{p}{)}
\PY{n}{md\PYZus{}}\PY{o}{=} \PY{n}{md\PYZus{}graph}\PY{o}{.}\PY{n}{edge\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{137}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{sys\PYZus{}db}\PY{o}{.}\PY{n}{delete\PYZus{}database}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{137}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
True
\end{Verbatim}
\end{tcolorbox}

    \hypertarget{data}{%
\section{Data}\label{data}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{139}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/results/}\PY{l+s+s2}{\PYZdq{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{o}{+}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ondemand}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ondemand}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{n}{f}\PY{p}{)}
    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
    \PY{n}{run\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{140}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/results/}\PY{l+s+s2}{\PYZdq{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{o}{+}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{performance}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{performance}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{n}{f}\PY{p}{)}
    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
    \PY{n}{run\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{141}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}data\PYZus{}measuments/parsec/ondemand}\PY{l+s+s2}{\PYZdq{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{p}{)}\PY{p}{:}
    \PY{k}{if} \PY{o+ow}{not} \PY{n}{f}\PY{o}{.}\PY{n}{endswith}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:} \PY{k}{continue}
    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{n}{f}\PY{p}{)}
    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
    \PY{n}{run\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \hypertarget{models}{%
\section{Models}\label{models}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{133}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/models/equation}\PY{l+s+s2}{\PYZdq{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{p}{)}\PY{p}{:}
    \PY{k}{if} \PY{o+ow}{not} \PY{n}{f}\PY{o}{.}\PY{n}{endswith}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:} \PY{k}{continue}
    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{n}{f}\PY{p}{)}
    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
    \PY{n}{aux}\PY{o}{=} \PY{n}{model\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
    \PY{n}{md\PYZus{}}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{p}{\PYZob{}}
        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}from}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}id}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}to}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{data}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{info}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{orig\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{vertex}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}key}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
    \PY{p}{\PYZcb{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{134}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/models/svr}\PY{l+s+s2}{\PYZdq{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{p}{)}\PY{p}{:}
    \PY{k}{if} \PY{o+ow}{not} \PY{n}{f}\PY{o}{.}\PY{n}{endswith}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:} \PY{k}{continue}
    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{n}{f}\PY{p}{)}
    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
    \PY{n}{aux}\PY{o}{=} \PY{n}{model\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
    \PY{n}{md\PYZus{}}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{p}{\PYZob{}}
        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}from}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}id}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}to}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{data}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{info}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{orig\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{vertex}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}key}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
    \PY{p}{\PYZcb{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \hypertarget{advantages}{%
\section{Advantages}\label{advantages}}

\begin{itemize}
\tightlist
\item
  No need to organize the files
\item
  Faster and easier queries
\item
  Shared data across programs and research
\end{itemize}

\hypertarget{disadvantages}{%
\section{Disadvantages}\label{disadvantages}}

\begin{itemize}
\tightlist
\item
  Need of db server
\item
  Need of network
\item
  Non local server maybe slow ( need to cache )
\end{itemize}

    \hypertarget{examples}{%
\section{Examples}\label{examples}}

\hypertarget{group-data-of-the-same-run}{%
\subsection{Group data of the same
run}\label{group-data-of-the-same-run}}

    \hypertarget{ondemand-groupby-argument}{%
\subsection{Ondemand groupby argument}\label{ondemand-groupby-argument}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{111}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{ondemand} \PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{aql}\PY{o}{.}\PY{n}{execute}\PY{p}{(}
\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
\PY{l+s+sd}{for d in run}
\PY{l+s+sd}{    filter d.config.governor == \PYZdq{}ondemand\PYZdq{}}
\PY{l+s+sd}{    filter d.config.data\PYZus{}descriptor.keys == [\PYZdq{}cores\PYZdq{}, \PYZdq{}input\PYZdq{}, \PYZdq{}repetitions\PYZdq{}]}
\PY{l+s+sd}{    collect args = d.config.arguments  into same\PYZus{}arg}
\PY{l+s+sd}{    return }
\PY{l+s+sd}{    \PYZob{}}
\PY{l+s+sd}{        keys: same\PYZus{}arg[*].d.\PYZus{}key,}
\PY{l+s+sd}{        pkgs: same\PYZus{}arg[*].d.config.pkg,}
\PY{l+s+sd}{        args}
\PY{l+s+sd}{    \PYZcb{}}
\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}\PY{p}{)}
\PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{ondemand}\PY{p}{:}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pkgs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{args}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
['./canneal']
['\_\_nt\_\_ 15000 2000 2500000.nets 128', '\_\_nt\_\_ 15000 2000 2500000.nets 256',
'\_\_nt\_\_ 15000 2000 2500000.nets 384', '\_\_nt\_\_ 15000 2000 2500000.nets 512',
'\_\_nt\_\_ 15000 2000 2500000.nets 640']
['./canneal', './canneal', './canneal']
['\_\_nt\_\_ 15000 2000 2500000.nets 3600', '\_\_nt\_\_ 15000 2000 2500000.nets 4200',
'\_\_nt\_\_ 15000 2000 2500000.nets 4800', '\_\_nt\_\_ 15000 2000 2500000.nets 5400',
'\_\_nt\_\_ 15000 2000 2500000.nets 6000']
['./fluidanimate', './fluidanimate']
['\_\_nt\_\_ 200 in\_500K.fluid out.fluid', '\_\_nt\_\_ 300 in\_500K.fluid out.fluid',
'\_\_nt\_\_ 400 in\_500K.fluid out.fluid', '\_\_nt\_\_ 500 in\_500K.fluid out.fluid',
'\_\_nt\_\_ 600 in\_500K.fluid out.fluid']
['./xhpl', './xhpl', './xhpl']
['\_\_nt\_\_ 7000', '\_\_nt\_\_ 8000', '\_\_nt\_\_ 9000', '\_\_nt\_\_ 10000', '\_\_nt\_\_ 11000']
['./blackscholes', './blackscholes', './blackscholes']
['\_\_nt\_\_ in\_5\_10.txt output.txt', '\_\_nt\_\_ in\_6\_10.txt output.txt', '\_\_nt\_\_
in\_7\_10.txt output.txt', '\_\_nt\_\_ in\_8\_10.txt output.txt', '\_\_nt\_\_ in\_9\_10.txt
output.txt']
['./x264', './x264', './x264']
['--quiet --qp 20 --partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid
--weightb --mixed-refs --no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4
--threads \_\_nt\_\_ -o eledream.264 eledream\_1920x1080\_306.y4m', '--quiet --qp 20
--partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs
--no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o
eledream.264 eledream\_1920x1080\_357.y4m', '--quiet --qp 20 --partitions
b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs --no-fast-
pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_408.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_459.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_512.y4m']
['./dedup', './dedup', './dedup', './dedup']
['-c -p -v -t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_05.iso -o output.dat.ddp', '-c -p -v
-t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_06.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
FC-6-x86\_64-disc1\_07.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
FC-6-x86\_64-disc1\_08.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
FC-6-x86\_64-disc1\_09.iso -o output.dat.ddp']
['./swaptions', './swaptions', './swaptions', './swaptions']
['-ns 32 -sm 2000000 -nt \_\_nt\_\_', '-ns 32 -sm 3000000 -nt \_\_nt\_\_', '-ns 32 -sm
4000000 -nt \_\_nt\_\_', '-ns 32 -sm 5000000 -nt \_\_nt\_\_', '-ns 32 -sm 6000000 -nt
\_\_nt\_\_']
['./ferret', './ferret', './ferret']
['corel\_5 lsh queries\_5 10 20 \_\_nt\_\_ output.txt', 'corel\_6 lsh queries\_6 10 20
\_\_nt\_\_ output.txt', 'corel\_7 lsh queries\_7 10 20 \_\_nt\_\_ output.txt', 'corel\_8
lsh queries\_8 10 20 \_\_nt\_\_ output.txt', 'corel\_9 lsh queries\_9 10 20 \_\_nt\_\_
output.txt']
['./vips', './vips', './vips']
['im\_benchmark orion\_10800x10800.v output.v', 'im\_benchmark orion\_12600x12600.v
output.v', 'im\_benchmark orion\_14400x14400.v output.v', 'im\_benchmark
orion\_16200x16200.v output.v', 'im\_benchmark orion\_18000x18000.v output.v']
['./openmc', './openmc', './openmc']
['input1', 'input2', 'input3', 'input4', 'input5']
['./bodytrack', './bodytrack', './bodytrack', './bodytrack']
['sequenceB\_1043 4 204 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 408 1000 10 0
\_\_nt\_\_', 'sequenceB\_1043 4 612 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 816 1000 10
0 \_\_nt\_\_', 'sequenceB\_1043 4 1020 1000 10 0 \_\_nt\_\_']
['./rtview', './rtview', './rtview']
['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 346 346',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 374 374',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 400 400',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 424 424',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 450 450']
['./freqmine', './freqmine', './freqmine']
['webdocs\_250k\_05.dat 11000', 'webdocs\_250k\_06.dat 11000', 'webdocs\_250k\_07.dat
11000', 'webdocs\_250k\_08.dat 11000', 'webdocs\_250k\_09.dat 11000']
    \end{Verbatim}

    \hypertarget{performance-groupby-argument}{%
\subsection{Performance groupby
argument}\label{performance-groupby-argument}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{112}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{performance} \PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{aql}\PY{o}{.}\PY{n}{execute}\PY{p}{(}
\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
\PY{l+s+sd}{for d in run}
\PY{l+s+sd}{    filter d.config.data\PYZus{}descriptor.extras.sensors}
\PY{l+s+sd}{    filter d.config.data\PYZus{}descriptor.keys == [\PYZdq{}cores\PYZdq{}, \PYZdq{}frequency\PYZdq{}, \PYZdq{}input\PYZdq{}, \PYZdq{}repetitions\PYZdq{}]}
\PY{l+s+sd}{    collect args = d.config.arguments  into same\PYZus{}arg}
\PY{l+s+sd}{    return }
\PY{l+s+sd}{    \PYZob{}}
\PY{l+s+sd}{        keys: same\PYZus{}arg[*].d.\PYZus{}key,}
\PY{l+s+sd}{        pkgs: same\PYZus{}arg[*].d.config.pkg,}
\PY{l+s+sd}{        args}
\PY{l+s+sd}{    \PYZcb{}}
\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}\PY{p}{)}
\PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{performance}\PY{p}{:}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pkgs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
    \PY{c+c1}{\PYZsh{}print(d[\PYZdq{}args\PYZdq{}])}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
['completo\_canneal\_1', 'completo\_canneal\_3', 'completo\_canneal\_4',
'completo\_canneal\_2']
['completo\_xhpl\_2', 'completo\_xhpl']
['completo\_black\_3', 'completo\_black\_2', 'completo\_black\_1']
['completo\_black\_4']
['completo\_black\_6', 'completo\_black\_5']
['completo\_x264\_4']
['completo\_x264\_2', 'completo\_x264\_1', 'completo\_x264\_3']
['completo\_dedup\_2', 'completo\_dedup\_1', 'completo\_dedup\_4', 'completo\_dedup\_3']
['completo\_swaptions\_3', 'completo\_swaptions\_2']
['completo\_swaptions\_1']
['completo\_fluid\_2']
['completo\_fluid\_4', 'completo\_fluid\_3']
['completo\_fluid\_1']
['completo\_ferret\_1', 'completo\_ferret\_2', 'completo\_ferret\_3',
'completo\_ferret\_4']
['completo\_vips\_2', 'completo\_vips\_3', 'completo\_vips\_4', 'completo\_vips\_1']
['completo\_openmc\_kernel\_novo']
['completo\_rtview\_1']
['completo\_bodytrack\_3']
['completo\_bodytrack\_2', 'completo\_bodytrack\_1']
['completo\_rtview\_2']
['completo\_rtview\_3']
['completo\_rtview\_4']
['completo\_freqmine\_2', 'completo\_freqmine\_1']
['completo\_freq']
    \end{Verbatim}

    \hypertarget{groupby-keys}{%
\subsection{Groupby keys}\label{groupby-keys}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{174}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{ts}\PY{o}{=} \PY{l+m+mf}{0.9}
\PY{n}{eq}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+s2}{equation}\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
\PY{n}{performance} \PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{aql}\PY{o}{.}\PY{n}{execute}\PY{p}{(}
\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}\PYZdq{}\PYZdq{}}
\PY{l+s+s2}{for d in model}
\PY{l+s+s2}{    filter d.info.type == }\PY{l+s+si}{\PYZob{}eq\PYZcb{}}
\PY{l+s+s2}{    filter d.train\PYZus{}sz == }\PY{l+s+si}{\PYZob{}ts\PYZcb{}}
\PY{l+s+s2}{    collect args = d.config.arguments  into same\PYZus{}arg}
\PY{l+s+s2}{    return }
\PY{l+s+s2}{    }\PY{l+s+s2}{\PYZob{}\PYZob{}}
\PY{l+s+s2}{        keys: same\PYZus{}arg[*].d.\PYZus{}key,}
\PY{l+s+s2}{        pkgs: same\PYZus{}arg[*].d.config.pkg,}
\PY{l+s+s2}{        maes: same\PYZus{}arg[*].d.info.mae,}
\PY{l+s+s2}{        args}
\PY{l+s+s2}{    \PYZcb{}\PYZcb{}}
\PY{l+s+s2}{\PYZdq{}\PYZdq{}\PYZdq{}}\PY{p}{)}
\PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{performance}\PY{p}{:}
    \PY{n}{idx\PYZus{}min}\PY{o}{=} \PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{maes}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{index}\PY{p}{(}\PY{n+nb}{min}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{maes}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pkgs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{idx\PYZus{}min}\PY{p}{]}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{args}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{keys}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{idx\PYZus{}min}\PY{p}{]}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{*}\PY{l+m+mi}{100}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
completo\_canneal\_2
['\_\_nt\_\_ 15000 2000 2500000.nets 128', '\_\_nt\_\_ 15000 2000 2500000.nets 256',
'\_\_nt\_\_ 15000 2000 2500000.nets 384', '\_\_nt\_\_ 15000 2000 2500000.nets 512',
'\_\_nt\_\_ 15000 2000 2500000.nets 640']
64562
--------------------------------------------------------------------------------
--------------------
completo\_xhpl
['\_\_nt\_\_ 7000', '\_\_nt\_\_ 8000', '\_\_nt\_\_ 9000', '\_\_nt\_\_ 10000', '\_\_nt\_\_ 11000']
64574
--------------------------------------------------------------------------------
--------------------
completo\_black\_1
['\_\_nt\_\_ in\_156K.txt out\_156K.txt', '\_\_nt\_\_ in\_312K.txt out\_312K.txt', '\_\_nt\_\_
in\_625K.txt out\_625K.txt', '\_\_nt\_\_ in\_1M.txt out\_1M.txt', '\_\_nt\_\_ in\_2M.txt
out\_2M.txt', '\_\_nt\_\_ in\_5M.txt out\_5M.txt', '\_\_nt\_\_ in\_10M.txt out\_10M.txt']
64592
--------------------------------------------------------------------------------
--------------------
completo\_black\_4
['\_\_nt\_\_ in\_312K.txt out\_312K.txt', '\_\_nt\_\_ in\_625K.txt out\_625K.txt', '\_\_nt\_\_
in\_1M.txt out\_1M.txt', '\_\_nt\_\_ in\_2M.txt out\_2M.txt', '\_\_nt\_\_ in\_5M.txt
out\_5M.txt', '\_\_nt\_\_ in\_10M.txt out\_10M.txt']
64603
--------------------------------------------------------------------------------
--------------------
completo\_black\_5
['\_\_nt\_\_ in\_5\_10.txt output.txt', '\_\_nt\_\_ in\_6\_10.txt output.txt', '\_\_nt\_\_
in\_7\_10.txt output.txt', '\_\_nt\_\_ in\_8\_10.txt output.txt', '\_\_nt\_\_ in\_9\_10.txt
output.txt']
64617
--------------------------------------------------------------------------------
--------------------
completo\_x264\_4
['--quiet --qp 20 --partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid
--weightb --mixed-refs --no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4
--threads \_\_nt\_\_ -o eledream.264 eledream\_1920x1080\_306.y4m', '--quiet --qp 20
--partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs
--no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o
eledream.264 eledream\_1920x1080\_357.y4m', '--quiet --qp 20 --partitions
b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs --no-fast-
pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_408.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_459.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_512.y4m']
64623
--------------------------------------------------------------------------------
--------------------
completo\_x264\_1
['--quiet --qp 20 --partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid
--weightb --mixed-refs --no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4
--threads \_\_nt\_\_ -o eledream.264 eledream\_1920x1080\_512.y4m', '--quiet --qp 20
--partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs
--no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o
eledream.264 eledream\_1920x1080\_459.y4m', '--quiet --qp 20 --partitions
b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs --no-fast-
pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_408.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_357.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
eledream\_1920x1080\_306.y4m']
64636
--------------------------------------------------------------------------------
--------------------
completo\_dedup\_3
['-c -p -v -t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_05.iso -o output.dat.ddp', '-c -p -v
-t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_06.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
FC-6-x86\_64-disc1\_07.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
FC-6-x86\_64-disc1\_08.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
FC-6-x86\_64-disc1\_09.iso -o output.dat.ddp']
64667
--------------------------------------------------------------------------------
--------------------
completo\_swaptions\_2
['-ns 32 -sm 2000000 -nt \_\_nt\_\_', '-ns 32 -sm 3000000 -nt \_\_nt\_\_', '-ns 32 -sm
4000000 -nt \_\_nt\_\_', '-ns 32 -sm 5000000 -nt \_\_nt\_\_', '-ns 32 -sm 6000000 -nt
\_\_nt\_\_']
64680
--------------------------------------------------------------------------------
--------------------
completo\_swaptions\_1
['-ns 64 -sm 2000000 -nt \_\_nt\_\_', '-ns 64 -sm 3000000 -nt \_\_nt\_\_', '-ns 64 -sm
4000000 -nt \_\_nt\_\_', '-ns 64 -sm 5000000 -nt \_\_nt\_\_', '-ns 64 -sm 6000000 -nt
\_\_nt\_\_', '-ns 64 -sm 7000000 -nt \_\_nt\_\_']
64686
--------------------------------------------------------------------------------
--------------------
completo\_fluid\_2
['1 125 in\_500K.fluid', '1 250 in\_500K.fluid', '1 500 in\_500K.fluid', '1 1000
in\_500K.fluid', '1 2000 in\_500K.fluid', '1 4000 in\_500K.fluid']
64697
--------------------------------------------------------------------------------
--------------------
completo\_fluid\_4
['1 200 in\_500K.fluid out.fluid', '1 300 in\_500K.fluid out.fluid', '1 400
in\_500K.fluid out.fluid', '1 500 in\_500K.fluid out.fluid', '1 600 in\_500K.fluid
out.fluid']
64703
--------------------------------------------------------------------------------
--------------------
completo\_fluid\_1
['1 62 in\_500K.fluid', '1 125 in\_500K.fluid', '1 250 in\_500K.fluid', '1 500
in\_500K.fluid', '1 1000 in\_500K.fluid', '1 2000 in\_500K.fluid', '1 4000
in\_500K.fluid']
64715
--------------------------------------------------------------------------------
--------------------
completo\_ferret\_1
['corel\_5 lsh queries\_5 10 20 \_\_nt\_\_ output.txt', 'corel\_6 lsh queries\_6 10 20
\_\_nt\_\_ output.txt', 'corel\_7 lsh queries\_7 10 20 \_\_nt\_\_ output.txt', 'corel\_8
lsh queries\_8 10 20 \_\_nt\_\_ output.txt', 'corel\_9 lsh queries\_9 10 20 \_\_nt\_\_
output.txt']
64721
--------------------------------------------------------------------------------
--------------------
completo\_vips\_4
['im\_benchmark orion\_10800x10800.v output.v', 'im\_benchmark orion\_12600x12600.v
output.v', 'im\_benchmark orion\_14400x14400.v output.v', 'im\_benchmark
orion\_16200x16200.v output.v', 'im\_benchmark orion\_18000x18000.v output.v']
64757
--------------------------------------------------------------------------------
--------------------
completo\_openmc\_kernel\_novo
['input1', 'input2', 'input3', 'input4', 'input5']
64769
--------------------------------------------------------------------------------
--------------------
completo\_rtview\_1
['nave.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 100 100', 'nave.obj
-automove -nthreads \_\_nt\_\_ -frames 10000 -res 140 140', 'nave.obj -automove
-nthreads \_\_nt\_\_ -frames 10000 -res 196 196', 'nave.obj -automove -nthreads
\_\_nt\_\_ -frames 10000 -res 274 274', 'nave.obj -automove -nthreads \_\_nt\_\_ -frames
10000 -res 384 384', 'nave.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 537
537']
64775
--------------------------------------------------------------------------------
--------------------
completo\_bodytrack\_3
['sequenceB\_1043 4 204 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 408 1000 10 0
\_\_nt\_\_', 'sequenceB\_1043 4 612 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 816 1000 10
0 \_\_nt\_\_', 'sequenceB\_1043 4 1020 1000 10 0 \_\_nt\_\_']
64781
--------------------------------------------------------------------------------
--------------------
completo\_bodytrack\_1
['sequenceB\_1043 4 64 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 128 1000 10 0
\_\_nt\_\_', 'sequenceB\_1043 4 256 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 512 1000 10
0 \_\_nt\_\_', 'sequenceB\_1043 4 1024 1000 10 0 \_\_nt\_\_']
64798
--------------------------------------------------------------------------------
--------------------
completo\_rtview\_2
['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 100 100',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 140 140',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 196 196',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 274 274',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 384 384',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 537 537']
64805
--------------------------------------------------------------------------------
--------------------
completo\_rtview\_3
['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 100 100',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 200 200',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 300 300',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 400 400',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 500 500']
64814
--------------------------------------------------------------------------------
--------------------
completo\_rtview\_4
['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 346 346',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 374 374',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 400 400',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 424 424',
'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 450 450']
64828
--------------------------------------------------------------------------------
--------------------
completo\_freqmine\_1
['webdocs\_250k\_05.dat 11000', 'webdocs\_250k\_06.dat 11000', 'webdocs\_250k\_07.dat
11000', 'webdocs\_250k\_08.dat 11000', 'webdocs\_250k\_09.dat 11000']
64849
--------------------------------------------------------------------------------
--------------------
completo\_freq
['webdocs.dat 90000']
64857
--------------------------------------------------------------------------------
--------------------
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{13}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{k+kn}{from} \PY{n+nn}{sklearn}\PY{n+nn}{.}\PY{n+nn}{model\PYZus{}selection} \PY{k+kn}{import} \PY{n}{train\PYZus{}test\PYZus{}split}

\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{rand}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{)}\PY{p}{)}

\PY{n}{X}\PY{o}{=} \PY{n}{df}\PY{p}{[}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{]}
\PY{n}{Y}\PY{o}{=} \PY{n}{df}\PY{p}{[}\PY{l+m+mi}{4}\PY{p}{]}

\PY{n}{Xtrain}\PY{p}{,} \PY{n}{\PYZus{}}\PY{p}{,} \PY{n}{Ytrain}\PY{p}{,} \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{train\PYZus{}test\PYZus{}split}\PY{p}{(}\PY{n}{X}\PY{p}{,} \PY{n}{Y}\PY{p}{,} \PY{n}{train\PYZus{}size}\PY{o}{=}\PY{l+m+mi}{5}\PY{p}{)}
\PY{n}{idx}\PY{o}{=} \PY{n}{Xtrain}\PY{o}{.}\PY{n}{index}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{19}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{df}\PY{o}{=}\PY{n}{df}\PY{o}{.}\PY{n}{sort\PYZus{}values}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
\PY{n}{df}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{19}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
          0         1         2         3         4         5         6  \textbackslash{}
9  0.777892  0.000535  0.239624  0.882285  0.951852  0.443976  0.966355
8  0.932123  0.027919  0.050847  0.592801  0.867408  0.164765  0.748370
3  0.675394  0.257285  0.523616  0.820933  0.657142  0.600372  0.220450
4  0.887332  0.288416  0.690859  0.118682  0.694805  0.611857  0.046453
5  0.800843  0.670505  0.759913  0.061861  0.054205  0.747787  0.822460
1  0.037602  0.672393  0.133586  0.004869  0.203755  0.168597  0.928038
2  0.702845  0.714748  0.298797  0.758865  0.256273  0.791469  0.854921
0  0.182818  0.803162  0.356512  0.286891  0.948721  0.509564  0.684132
7  0.475779  0.818054  0.397074  0.239534  0.630659  0.024413  0.467922
6  0.539352  0.862582  0.722857  0.116203  0.177171  0.788025  0.039456

          7         8         9
9  0.035550  0.404241  0.547084
8  0.857539  0.840110  0.636682
3  0.132569  0.576291  0.461130
4  0.279059  0.129637  0.241594
5  0.924498  0.617560  0.317410
1  0.621204  0.808938  0.369552
2  0.004346  0.345329  0.451903
0  0.576865  0.166287  0.171629
7  0.421451  0.344078  0.591117
6  0.796943  0.106906  0.065239
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{20}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{f}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}json}\PY{p}{(}\PY{p}{)}
\PY{n}{df2}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{read\PYZus{}json}\PY{p}{(}\PY{n}{f}\PY{p}{)}
\PY{n}{df2}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{20}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
          0         1         2         3         4         5         6  \textbackslash{}
9  0.777892  0.000535  0.239624  0.882285  0.951852  0.443976  0.966355
8  0.932123  0.027919  0.050847  0.592801  0.867408  0.164765  0.748370
3  0.675394  0.257285  0.523616  0.820933  0.657142  0.600372  0.220450
4  0.887332  0.288416  0.690859  0.118682  0.694805  0.611857  0.046453
5  0.800843  0.670505  0.759913  0.061861  0.054205  0.747787  0.822460
1  0.037602  0.672393  0.133586  0.004869  0.203755  0.168597  0.928038
2  0.702845  0.714748  0.298797  0.758865  0.256273  0.791469  0.854921
0  0.182818  0.803162  0.356512  0.286891  0.948721  0.509564  0.684132
7  0.475779  0.818054  0.397074  0.239534  0.630659  0.024413  0.467922
6  0.539352  0.862582  0.722857  0.116203  0.177171  0.788025  0.039456

          7         8         9
9  0.035550  0.404241  0.547084
8  0.857539  0.840110  0.636682
3  0.132569  0.576291  0.461130
4  0.279059  0.129637  0.241594
5  0.924498  0.617560  0.317410
1  0.621204  0.808938  0.369552
2  0.004346  0.345329  0.451903
0  0.576865  0.166287  0.171629
7  0.421451  0.344078  0.591117
6  0.796943  0.106906  0.065239
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{28}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{x}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{def}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{AAA}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fff}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{CCC}\PY{l+s+s2}{\PYZdq{}}\PY{p}{\PYZcb{}}
\PY{n}{y}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{abc\PYZus{}def}\PY{l+s+s2}{\PYZdq{}}

\PY{n}{title}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
\PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{x}\PY{p}{:}
    \PY{k}{if} \PY{n}{k} \PY{o+ow}{in} \PY{n}{y}\PY{p}{:}
        \PY{n}{title}\PY{o}{=} \PY{n}{x}\PY{p}{[}\PY{n}{k}\PY{p}{]}
\PY{n}{title}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{28}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
'AAA'
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{44}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{l+m+mf}{46.615389}\PY{o}{\PYZhy{}}\PY{l+m+mf}{59.829612}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{44}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
-13.214222999999997
\end{Verbatim}
\end{tcolorbox}

    % Add a bibliography block to the postdoc

\end{document}