### 2022-04-05 thesis/main.tex
umons template

\documentclass[
papersize=a4,
pagelayout=default,
fontname=latinmodern,
fontsize=11pt,
twoside,
final,
faculty=fpms
]{umons-Thesis}

% Styles
\headstyles{umons}
\pagestyle{umons}

\usepackage[main=english]{babel}
\usepackage[autostyle=true]{csquotes}
\usepackage[babel=true]{microtype}
\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{siunitx}
\sisetup{%
	binary-units = true,%	loads units for binary data (\bit, \byte)
}
\usepackage[%
hidelinks,%				hides links (removing color and border)
pdfusetitle,%			uses the content of \author{}... for metadata
pdfdisplaydoctitle,%	displays document title instead of file name in title bar
]{hyperref}
\usepackage{cleveref}
\usepackage{outlines}
\usepackage{listings}

\lstset { %
	language=C++,
	backgroundcolor=\color{black!5}, % set backgroundcolor
	basicstyle=\footnotesize,% basic font setting
}

\newcommand*\rot{\rotatebox{90}}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\DeclareOldFontCommand{\rm}{\normalfont\rmfamily}{\mathrm}

%%        %%
% DOCUMENT %
% ======== %

% Metadata
% --------

\author{Vitor \textsc{Ramos Gomes da Silva}}

\date{\today}

\title{Title}
%\title{My very long two-line PhD thesis title}
%\title{My supra long PhD thesis title just to verify that nothing is broken because of a long title}

\committeeMembers{%
	Prof. Carlos Alberto \textsc{VALDERRAMA}, Supervisor\\
	Prof. Pierre \textsc{MANEBACK}, Co-supervisor\\
	Prof. Thierry \textsc{DUTOIT}, Chair\\
	Prof. Sidi \textsc{MAHMOUDI}, Chair\\
	Prof. Samuel \textsc{XAVIER-DE-SOUZA}, Chair 
}

% Text
% ----

\begin{document}

	\umonsThesisTitlePage

	\frontmatter

	\begin{umonsThesisAbstract}
		Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
		Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have priority information on the application, which leads to inefficient software operation. This work proposes an analytical model that can be used to estimate energy-optimal software operation. 
		This work proposes a new energy model that can be used as a basis to optimize DVFS and DPM, as well as to assess the contribution of parameters such as parallelism level and dynamic power to the energy consumption.
		Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
	\end{umonsThesisAbstract}

	% \vspace{2cm}
	% \textbf{Keywords :} Energy Model; Dynamic Frequency and Voltage Scaling; Dynamic Power Management; High Performance Computing

	\tableofcontents*
	\listoffigures*
	\listoftables*

	\mainmatter

	\chapter{Introduction}

	\include{intro/intro}

	\chapter{Proposed Models}
	In this Section, the proposed power, performance and energy models are derived and validated
	\include{models/models}
	\include{phases/phases}
	\include{optmizers/optmizers}

	\chapter{Framework}
	This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.
	\include{fingerprint/fingerprint}
	\include{pascalanalyzer/pascalanalyzer}

	\chapter{Results (model validation)}
	This chapter will describe all the experiments that validate the models constraints.
	\include{experiments/experiments}

	\chapter{Conclusions}

	\bibliographystyle{plain}
	\bibliography{references}

	\backmatter

\end{document}