### 2022-05-03 thesis/phases/phases.tex
phases

\section{Model with phases} \label{sec:model_with_phases}

Dynamic frequency scaling and dynamic power management are the leading software techniques for energy saving in modern computers. These techniques work through dynamic workload planning, and switching from one power state to another is crucial for them. This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques based on the time-division and allocated resources. We propose a methodology combining measurement data with a simulation to provide insights into choosing the best time intervals. Furthermore, we evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques identifying that generally exists a limit to the number of divisions that the algorithm can benefit from, giving a lower bound to the minimal time interval.

\section{Introduction to models on time} \label{sec:introduction_to_models_on_time}
Energy efficiency in computing has always been a concern, especially in high-performance computing (HPC), where a small percentage in energy savings can significantly reduce maintenance costs or environmental impact due to its scale of consumption. For example, the leading Petaflop supercomputers consume are in a range of 1–18 MW of electrical power, with 1.5 MW on average, which we can estimate to be into a millions dollars range of electricity bills annually \cite{Group2012HandbookSahni}. Also,  the International Energy Agency  \cite{iea_2021} estimated that global data center electricity use in 2020 was around 200-250 TWh, or around 1\% of the global electricity demand, which can generate as much pollution as a nation like Argentina \cite{Mathew2012Energy-awareNetworks}.

Given the importance of this topic, modern processors of desktop, mobile phones, and HPC  have internal power management mechanisms, and that is because the processor is one of the main components that drain energy and can reach 50\% of the total system consumption \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}. Within the mechanisms implemented by the processors, the main controllable via software are Dynamic Frequency and Voltage Scaling (DVFS) and Dynamic Power Management (DPM) ~\cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}.

DPM encompasses a set of techniques that take advantage of different power levels on the system, such as active, idle, and disabled~\cite{CARDOSO201717, Shuja2012Energy-efficientCenters, Benini2000AManagement}. The basic idea is to shut down devices and wake them up when necessary. However, this is not trivial, that are several variables that need to take into account as the switching cost, in terms of time and energy, for how long it will remain in that state,  the power consumption on every state, and a lot more.

The DPM technique is quite interesting on systems where static power is very high or where the system is idle for a long time. In these cases, some works report savings of up to 70\% in energy~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.

DVFS allows real-time control of the frequency and voltage depending on the needs. This method is motivated by the fact that we can approximate the relationship of the frequency with the power as a cubic equation and frequency and performance as a linear equation \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}, implying that a reduction in frequency has a cubic impact on power and linear on performance. Moreover, reducing the frequency leads to a longer execution time, increasing energy consumption. As a result, determining the best voltage and frequency to employ in all conditions is a challenging task.

\section{Phases related work} \label{sec:related_work_phases}
Some related work

\section{Optimal energy configuration} \label{sec:Optimal_energy_configuration}
Splitting a program into phases, or time intervals with a selected set of execution parameters, has an impact on energy consumption. To find the ideal phase division, we need to compute the optimal energy configuration for a given number and duration of phases. For that, many different approaches could be applied, each having different trade-offs.

Using models ref [x], or a static analysis of the application, allows obtaining a very fast but less precise estimate. On the other hand, using the application and evaluating its behavior in each possible context would be a more reliable but too expensive method. Indeed, since the number of combinations grows exponentially, this is a very time consuming approach. Instead, our analysis approach will provide both, speed and precision. We will simply run the target application and collect performance data from every possible execution context. By context, we can consider, for instance, the frequency and/or number of cores. This data will then be used to analyse the behavior of phases in all possible combinations. For this, we assume that the number of instructions of the executed program is the same regardless of frequency and number of threads.

\begin{equation}
    I=c_{k1}f_1T_1=c_{k2}f_2T_2
\end{equation}

where $I$ is the number of instructions, $c_{k1}$ the rate of instructions per cycle, $f$ the frequency of the processor and $T$ the execution time. All the required parameters can be collected by simply running the application with the available frequencies.

From this equation we can split the execution of the program into several different phases $i$ as follows:
% \begin{equation}
%     I=xI+(1-x)I=xc_{k1}f_1+(1-x)c_{k2}f_2
% \end{equation}
\begin{equation}
    I=\sum_{i}{a_iI} %=\sum_i{a_ic_{ki}f_i}
\end{equation}

with $a_i$ as the percentages of instructions executed with frequency $f_i$, we can compute any combination of phases. For convenience, we write this as a function of the time:
% \begin{equation}
% T=\frac{I}{c_{k1}f_1}
% \end{equation}
\begin{equation}
I=Tc_{k}f
\end{equation}
\begin{equation}
    I=\sum_{i}{a_iI}=\sum_i{a_iTc_{ki}f_i}
\end{equation}
% \begin{equation}
% I=xI+(1-x)I=xTc_{k1}f_1+(1-x)Tc_{k2}f_2
% \end{equation}

the application being divided into phases of different duration, $a_i$  becomes the duration of the phase as a percentage of the total execution time. This general equation does not limit the number and duration of phases, and several phases can have the same frequency.

In the event that we decide to include the number of cores as a context parameter, the equation becomes more complex. Indeed, the number of cores can completely change the application's behavior but, depending on the type of application, it's still possible to find similarities. Many HPC applications use data parallelism, and the Amdahl's law can model most of it. For this kind of applications, we can estimate that:
\begin{equation}
T=\frac{T_s}{S}
\end{equation}
\begin{equation}
T_s=\frac{I}{c_kf}
\end{equation}
\begin{equation}
S=\frac{1}{1-w+\frac{w}{p}}
\end{equation}
\begin{equation}
T=\frac{I(1-w+\frac{w}{p})}{c_kf}=\frac{I}{c_kf}-\frac{Iw}{c_kf}+\frac{Iw}{c_kfp}
\end{equation}

Looking at a fixed frequency, we arrive at a very similar structure as before:
\begin{equation}
T=c_1I+\frac{c_2I}{p}
\end{equation}

As previously, we can use $I$ as the number of reference instructions, and compute $a$ as the percentage of single and multi-core instructions, or express it as a percentage of the total time T:
\begin{equation}
aT=c_1aI+\frac{c_2aI}{p}
\end{equation}

The extent of the hypotheses will later be validated, however, by taking advantage of both principles, we can now, based on real measurements, estimate any phase division and possible context combination.

% Here a need to show that the time is proportional to the number of executed instructions.
% WHY ?
% I am trying to say here that we can split the program using a percentage of the time because the time is proportional to the number of executed instructions. If we complete x% of the time, that will be x% of instructions. This implies that we can resume and continue the program from any point ensuring that we executed all instructions. If we have a previous execution for each configuration, we can compute any combination with minimal error.

\section{Data collection} \label{sec:data_collection}
Data, power samples, can be gathered by running the application in several possible context configurations. In a system with 32 cores that can run at 13 frequencies, we can have 416 (32*13) possible configurations. Power samples can be taken from dedicated sensors in the IPMI system [REF]. To avoid overload, a sample rate of 0.5 seconds is more than sufficient. Additionally, the behavior of an application can be affected by the workload, so it is also interesting to run applications with different inputs and assess their impact on the collected data [REF victor alex paper].

% We can consider different input sizes as an new application...

\section{The simulator} \label{sec:the_simulator}
To analyse the impact of configurations and identify optimal time-intervals, phases, we propose an integrator, which computes the energy on an interval for a given configuration, and a reference, a brute force test of all possible combinations. The integrator has order zero, and computers the energy between two timestamps, as shown in the figure \ref{fig:zero_order}, where the red and yellow parts area are the energy value.

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{phases/figures/zero_order.png}
    \caption{Power vs time intervals for a given application. The energy associated area is colored.}
    \label{fig:zero_order}
\end{figure}

Power and energy can be computed for a given time interval based on power samples. The pseudo-algorithm for this computation is described bellow:

\begin{lstlisting}
timeslice_energy(info, start, stop)
{
   power = info.power;
   time = info.time;

   t1 = start_time + (elapsed_time) * start;
   t2 = start_time + (elapsed_time) * stop;
   en = 0;

   // first time that is greater than t1
   for(i=0; i<nsamples; i++)
       if(time[i] >= t1)
           break;
   v1 = (i == 0)?0:i-1;
   // first time that is greater than t2
   for(i=v1; i<nsamples; i++)
       if(time[i] >= t2)
           break;
   v2 = (i == 0)?0:i-1;

   // interpolate and integrates the power samples
   if(v1 == v2)
       return (t2-t1)*power[v1];

   en = (time[v1+1]-t1)*power[v1];
   en += (t2-time[v2])*power[v2];
   // integrate samples in between
   for(i=v1+1; i<v2; i++)
       en += (time[i+1]-time[i])*power[i];
   return en;
}
\end{lstlisting}

The reference can simply be obtained by applying brute-force, which cycles through all configurations for all divisions and compute the minimal energy. Other more effective, but less accurate, methods can be applied.

\begin{lstlisting}
bruteforce(data, phases)
{
   min_en = -1;
   total_en = 0;
   for (i = 0; i < nphases; i++)
   {
       min_en = -1;
       for (j = 0; j<data.nconfigs; j++)
       {
           en = timeslice_energy(data.info[j], 
                        phases[i], phases[i + 1]);
           if (min_en == -1 || en < min_en)
           {
               min_en = en;
               min_conf = data.config[j];
           }
       }
       total_en += min_en;
   }
   return total_en;
}
\end{lstlisting}

\section{Optimizing phases} \label{sec:optimizing_phases}
The main goal of this section is to identify the optimal number and duration of time intervals. Indeed, several short phases can have an impact on switching overhead as well as few long phases on the overall energy.

Using the objective function, brute force can identify the optimal energy on each phase and any minimization method can provide the best phase division. By computing in a range of a number of phases, we can find the one that brings the lowest energy value. A genetic algorithm GA can be used to analyse a range of, for instance, 3 to 100 phases. A GA for a population of 10K individuals with a 0.1 elitism will ensure that 10\% of the best individuals will survive and 0.1 mutations will avoid local minimums within a maximum of 300 steps. A reproduction function will combine the half-phases of one individual with another, and the mutation will randomly change a division.

\section{Phases experiments results} \label{sec:results_phases_experiments}

% Normalization
% Show that number of instructions is the same...

In order to validate the approach, HPC benchmarks where chosen, 13 applications from PARSEC 3.0, HPC, and Openmc.

\begin{figure}[H]
\includegraphics[width=\columnwidth]{fingerprint/figures/energy_per_phase.png}
    \caption{Relative energy vs number of phases using 13 applications from PARSEC 3.0, HPC, and Openmc benchmarks. Energy compared to the single-phase optimal configuration.}
    \label{fig:relative_energy}
\end{figure}

This evaluation shows very interesting results, as shown in the figure \ref{fig:relative_energy}. Indeed, for all cases, an optimal single-phase configuration gives the highest energy draw. On the contrary, it is not necessary to split the application into more than 35 phases. On average, the energy consumption stood still after 35 phases.

The results attend our expectation that the optimal with n divisions will always be better or equal than n-1 divisions. This can be explained by the fact that after we find the ideal frequency and core curves, increasing the number of phases will not change the control signals. Therefore the energy will remain the same, and there is no reason to increase the number of phases beyond that actually.

It is important to note that the additional energy savings that theoretically may be achieved when running with a greater number of phases will have a negative impact on performance and context switching. Moreover, the maximal energy saving compared to the single-phase case was just 2.6\%. This may be due to the type of applications, where a fine-grain fast control will not save more energy. On the contrary, this switching will cause to consume a lot more energy, and this is likely to happen with online DVFS algorithms, very sensitive to fast workload variations [REF].

We evaluated separately the impact on energy when varying the number of cores. Energy-cores figures doesn't change fundamentally varying the number of phases. Indeed, bellow we can see some examples with optimal energy cores selections for 60 and 70 phases (figure \ref{fig:cores_control}). Note the very small differences that will not change significantly the energy figures.

\begin{figure}[h!]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_cores_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_cores_signals_cmp.png}
%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
\caption{Energy and number of cores with 37 and 99 phases.}
    \label{fig:cores_control}
\end{figure}

We observed a similar behavior just considering frequency control energy optimisation (figure \ref{fig:freq_control}).

\begin{figure}[h!]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_freq_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_freq_signals_cmp.png}
%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
\caption{Energy and frequency with 37 and 99 phases.}
    \label{fig:freq_control}
\end{figure}

The images below show the phases for all applications:
\begin{figure}[H]
\includegraphics[width=\columnwidth]{phases/figures/phase_division_35.png}
\end{figure}
In most applications, we can see a pattern of 3 phases wich for this applications are generally, and setup phase where it loads data from the disk, a computation phase where the data is processed, and a finalization phase where data is written back to the disk or display to the user.

\begin{figure}[H]
\includegraphics[width=\columnwidth]{phases/figures/phase_division_3.png}
\end{figure}
Plotting for 3 phases, we can observe that better.